Verification of Stencil Computations through Spatial Specifications

### 2016-07-06 paper.tex
reformat
typo
tweak

We define a semantic model for our stencil specification language.
This model has a number of purposes: (1) it is used in the inference,
checking, and program synthesis algorithms (introduced in
Section~\ref{sec:analysis}); (2) gives a basis for the equational
theory and approximations of the previous section; and (3) can be used
to guide correct implementations. The model is a denotational
semantics, in terms of sets of vectors which we call \emph{index
  schemes}. We introduce first the notion of indexing schemes
and \emph{unification} which gives the domain of our model
and forms the basis of the rest of the algorithsm for our
specificaiton. We then define the denotational model

@@

\begin{definition}[Schematic]
The partial function $\textsf{schematic}$ maps 
source-language indexing terms $(e_1, \ldots, e_n)$
into $n$-dimensional indexing schemes 
($\in \domainVal^n$), defined:
%
\begin{align*}
\mathsf{schematic}(e_1, \ldots, e_n) & = 
[\downarrow\!e_1 \ldots \downarrow\!e_n ] 
\quad \textit{iff} \; \forall i . \downarrow\!e_i \neq \bot 
\end{align*}
%
The $\downarrow$ operator is a partial\footnote{We represent
  partiality explicitly in our types with the set $A + \bot$.} function 
 $\downarrow : \textit{expr} \rightarrow (\domainVal + \bot)$
which maps each individal indexing expression to a component of
the indexing scheme, defined: 
%%
\begin{align*}
\downarrow\!e
 & =  \begin{cases}
a & e \equiv i + a \; \wedge \; \mathsf{neigh}(i + a) \\
-a & e \equiv i - a \; \wedge \; \mathsf{neigh}(i - a) \\
\infty & \textit{$\mathsf{IV}(e) = \emptyset$} \\
\bot   & \textit{otherwise}
\end{cases}
\end{align*}
%%
where $\mathsf{IV}(e)$ is the set of free induction variables
appearing in an expression, and $\neigh{}$ classifies expressions which 
are constant offsets from an induction variable
(Defn.~\ref{def:neighbour}). In the third case, 
if the expression is not a neighbourhood index but is not
a relative index either (\ie{} is not defined in terms
of an induction variable) then it is mapped to $\infty$, otherwise
$\downarrow\!e$ is undefined. The $\mathsf{schematic}$ function
is undefined if $\downarrow$ of any of its components $e_i$ is
undefined.

@@

The pairwise permutation $u \bowtie v$ takes two vectors and builds the matrix
where the rows are all possible $n$-vectors generated by
non-deterministic picking each $i^{th}$ entry from either the
$i^{th}$ entry in $u$ or the $i^{th}$ entry in $j$. For example:
%
\begin{equation*}
\vthreeh{0}{1}{2} \bowtie \vthreeh{3}{4}{5} =
\setlength{\arraycolsep}{0.5em}
\left[\begin{array}{ccc}
0 & 1 & 2 \\
0 & 1 & 5 \\
0 & 4 & 2 \\
0 & 4 & 5 \\
3 & 1 & 2 \\
3 & 1 & 5 \\
3 & 4 & 2 \\
3 & 4 & 5
\end{array}
\right]
\end{equation*}
%
The $2^n$ unique choices for $\bowtie$ on $n$-vectors
corresponds to taking all bit-strings of length $n$ and
selecting from $u$ for $1$ and $v$ for $0$. Thus, we can
defined the pairwise permutation matrix as:
%
\begin{align*}
(u \bowtie v)_{i,j} & = b_{i,j} u_j + \neg b_{i,j} v_j
\end{align*}
%
where $b$ is a logical matrix where $b_{i,j}$ is the $j$-th bit of the integer
$i - 1$.
%%
\end{definition}

>>>>

We define a semantic model for our stencil specification language.
This model has a number of purposes: (1) it is used in the inference,
checking, and program synthesis algorithms (introduced in
Section~\ref{sec:analysis}); (2) gives a basis for the equational
theory and approximations of the previous section; and (3) can be used
to guide correct implementations. The model is a denotational
semantics, in terms of sets of vectors which we call \emph{index
  schemes}. We introduce first the notion of indexing schemes
and \emph{unification} which gives the domain of our model
and forms the basis of the rest of the algorithsm for our
specificaiton. We then define the denotational model.

@@

\begin{definition}[Schematic]
The partial function $\textsf{schematic}$ maps 
source-language indexing terms $(e_1, \ldots, e_n)$
into $n$-dimensional index schemes 
$\in (\domainVal)^n$, defined:
%
\begin{align*}
\mathsf{schematic}(e_1, \ldots, e_n) & = 
[\downarrow\!e_1 \ldots \downarrow\!e_n ] 
\quad \textit{iff} \; \forall i . \downarrow\!e_i \neq \bot 
\end{align*}
%
The $\downarrow$ operator is a partial\footnote{We represent
  partiality explicitly in our types with the set $A + \bot$.} function 
 $\downarrow : \textit{expr} \rightarrow (\domainVal + \bot)$
which maps each individal indexing expression to a component of
the indexing scheme, defined: 
%%
\begin{align*}
\downarrow\!e
 & =  \begin{cases}
a & e \equiv i + a \; \wedge \; \mathsf{neigh}(i + a) \\
-a & e \equiv i - a \; \wedge \; \mathsf{neigh}(i - a) \\
\infty & \textit{$\mathsf{IV}(e) = \emptyset$} \\
\bot   & \textit{otherwise}
\end{cases}
\end{align*}
%%
where $\mathsf{IV}(e)$ is the set of free induction variables
appearing in an expression, and $\neigh{}$ classifies expressions which 
are constant offsets from an induction variable
(Defn.~\ref{def:neighbour}). In the third case, 
if the expression is not a neighbourhood index but is not
a relative index either (\ie{} is not defined in terms
of an induction variable) then it is mapped to $\infty$, otherwise
$\downarrow\!e$ is undefined. The $\mathsf{schematic}$ function
is undefined if $\downarrow$ of any of its components $e_i$ is
undefined.

@@

The pairwise permutation $u \bowtie v$ takes two vectors and builds the matrix
where the rows are all possible $n$-vectors generated by
non-deterministic picking each $i^{th}$ entry from either the
$i^{th}$ entry in $u$ or the $i^{th}$ entry in $j$. For example:
%
\begin{equation*}
\vthreeh{0}{1}{2} \bowtie \vthreeh{3}{4}{5} =
\left[
\setlength{\arraycolsep}{0.42em}
\begin{array}{cccccccc}
0 & 0 & 0 & 0 & 3 & 3 & 3 & 3 \\ 
1 & 1 & 4 & 4 & 1 & 1 & 4 & 4 \\
2 & 5 & 2 & 5 & 2 & 5 & 2 & 5
\end{array}
\right]
% \\
%\setlength{\arraycolsep}{0.5em}
%\left[\begin{array}{ccc}
%0 & 1 & 2 \\
%0 & 1 & 5 \\
%0 & 4 & 2 \\
%0 & 4 & 5 \\
%3 & 1 & 2 \\
%3 & 1 & 5 \\
%3 & 4 & 2 \\
%3 & 4 & 5
%\end{array}
%\right]
\end{equation*}
%
The $2^n$ unique choices for $\bowtie$ on $n$-vectors
corresponds to taking all bit-strings of length $n$ and
selecting from $u$ for $1$ and $v$ for $0$. Thus, we can
defined the pairwise permutation matrix as:
%
\begin{align*}
(u \bowtie v)_{i,j} & = b_{i,j} u_j + \neg b_{i,j} v_j
\end{align*}
%
where $b$ is a logical matrix where $b_{i,j}$ is the $j$-th bit of the integer
$i - 1$.
%%
\dnote{I didn't understand mistral's description of unification
wrt to $\tensor$. I can't work it out either! Mistral have another
go given my above explanation}
\end{definition}
